---
title: "class 12 pt 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB STATS 
Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. 
```{r}
# read in file 
stats <- read.csv("Data Export Summary.csv", row.names=1)
# calculate percentage values asked for each 'method' 
ans <- states$Total / sum(states$Total) *100
names(ans) <- rownames(stats)
round(ans, 2)
```


Q2. Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.
```{r}
round(totalproteins / totalsum *100, 2)
```

Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?



#Read a single PDB structure into R
```{r}
library(bio3d)
pbd <- read.pdb("1HSG")
pbd
```
## Select protein only atoms 
here use `atom.select()` function to do this 
```{r}
prot <- atom.select(pdb, "protein", value = TRUE)
lig <- atom.select(pdb, "ligand", value = TRUE)
```
write out these new PDB objects with the `write.pbd()` function
```{r}
write.pdb(lig, file="1HSG_ligand.pbd")
write.pdb(prot, file = "1HSG_protein.pbd")
```

```{r}
attributes(pdb)
```

```{r}
pdb$atom
```

