---
title: "mini proj"
author: "Rachael Hall"
date: "2/6/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
first; import/read data 
```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```
second; weird tings going on that we will ignore xd including the first and second ID and diagnosis columes and col 33 
```{r}
wisc.data <- as.matrix((wisc.df[,3:32]))
head(wisc.data)
```
```{r}
nrow(wisc.data)
```
```{r}
table(wisc.df$diagnosis)
```
```{r}

grep("_mean", colnames(wisc.data), value = TRUE)
```
```{r}
length(grep("_mean", colnames(wisc.data)))
```



#PRINCIPAL COMPONENT ANALYSIS wi
It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

The input variables use different units of measurement.

- The input variables have significantly different variances.
- Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the `colMeans()` and `apply()` functions like you’ve done before.

```{r}
round(apply(wisc.data, 2, sd), 2)
```
set scale = tRUE 

```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```

the first PC captures 44.27% of the original variance. Note that 72.6% is captured in the frst 3 PCs 

```{r}
biplot(wisc.pr)
```
 
this aint it. do ur own and color by diagnosis! 

```{r}
attributes(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1],wisc.pr$x[,2], col=wisc.df$diagnosis)
abline(col="gray", thy=2),
```
```{r}
#cluster og data?
  hclust(dist(wisc.data))
```
combine results of PCA with clustering 
```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:3]), method = "ward.D2")
plot(wisc.pr.hclust)
abline(h=40)
```

to make the tree looook not ugly we need to cut it with `cutree` function 
```{r}
grps3 <- cutree(wisc.pr.hclust, k=2)
table(grps3)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps3)
```

use `table()` function to compare the $diagnosis vector with our cluster results vector 
```{r}
table(grps3, wisc.df$diagnosis)
```

```{r}
new <- read.csv("new_samples.csv")
new
```
use predict func with out previous pCA model 
```{r}
npc <- predict(wisc.pr, newdata = new)
npc
```
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=15, cex=3)
text(npc[,1], npc[,2], labels = c(1,2), col = "white")
```

